---
title: "Time Series Analysis with forecast package in R"
output: pdf_document
author: The Data Science Show
---

Youtube tutorial [Time Series Analysis with forecast Package in R Example Tutorial - The Data Science Show](https://www.youtube.com/watch?v=Zu1iimmsKD0). 

# Time Series Data
Data that changes over time, e.g., stock price, sales growth.

# AR, MA & ARIMA Models

* Autoregressive (AR) Model

$\gamma_t = \beta_1\gamma_{_t-_1} + \beta_2\gamma_{_t-_2}+...+\beta_\rho\gamma_{_t-_\rho}$

* Moving Average (MA) Model

$\gamma_t = \epsilon_t+\alpha_1\epsilon_{_t-_1}+\alpha_2\epsilon_{_t-_2}+...+\alpha_\delta\epsilon_{_t-_\delta}$

* ARMA

$\gamma_t = \beta_1\gamma_{_t-1}+\beta_2\gamma_{_t-2}+\beta_3\gamma_{_t-3}+...+\beta_\rho\gamma_{_t-\rho}+\epsilon_t+\alpha_1\epsilon_{_t-1}+\alpha_2\epsilon_{_t-2}+...+\alpha_\delta\epsilon_{_t-\delta}$

```{r}
install.packages("WDI")
library("WDI")
gdp <- WDI(country=c("US"), indicator=c("NY.GDP.PCAP.CD"), start=1960, end=2016)
head(gdp)
```
The column names of the dataset are a little clunky so let's fix this with the names() function...
```{r}
names(gdp) <- c("iso2c", "country", "GDPPerCap", "year")
head(gdp)
```

We'll now organise the dataset by year (ASC) order...
```{r}
gdp <- gdp[order(gdp$year),] #remember comma!!
head(gdp)
```

An initial plot shows GDP Per Capita trend (if any)
```{r}
plot(GDPPerCap~year, data = gdp)
```
Let's make things a little more eye catching using ggplot...
```{r}
library(ggplot2)
ggplot(gdp, aes(x=gdp$year, y=gdp$GDPPerCap)) +
        geom_point(
                color="orange",
                fill="#69b3a2",
                alpha=0.5,
                size=2,
                shape=21,
                stroke=1
        )
```



