---
title: "Time Series Analysis with forecast package in R"
output: pdf_document
author: The Data Science Show
---

Youtube tutorial [Time Series Analysis with forecast Package in R Example Tutorial - The Data Science Show](https://www.youtube.com/watch?v=Zu1iimmsKD0). 

# Time Series Data
Data that changes over time, e.g., stock price, sales growth.

# AR, MA & ARIMA Models

* Autoregressive (AR) Model

$\gamma_t = \beta_1\gamma_{_t-_1} + \beta_2\gamma_{_t-_2}+...+\beta_\rho\gamma_{_t-_\rho}$

* Moving Average (MA) Model

$\gamma_t = \epsilon_t+\alpha_1\epsilon_{_t-_1}+\alpha_2\epsilon_{_t-_2}+...+\alpha_\delta\epsilon_{_t-_\delta}$

* ARMA

$\gamma_t = \beta_1\gamma_{_t-1}+\beta_2\gamma_{_t-2}+\beta_3\gamma_{_t-3}+...+\beta_\rho\gamma_{_t-\rho}+\epsilon_t+\alpha_1\epsilon_{_t-1}+\alpha_2\epsilon_{_t-2}+...+\alpha_\delta\epsilon_{_t-\delta}$

```{r obtain-data}
# install.packages("WDI")
library("WDI")
gdp <- WDI(country=c("US"), indicator=c("NY.GDP.PCAP.CD"), start=1960, end=2016)
head(gdp)
```
The column names of the dataset are a little clunky so let's fix this with the names() function...
```{r change-colnames}
names(gdp) <- c("iso2c", "country", "GDPPerCap", "year")
head(gdp)
```

We'll now organise the dataset by year (ASC) order...
```{r reorder-data-by-year(ASC)}
gdp <- gdp[order(gdp$year),] #remember comma!!
head(gdp)
```

An initial plot shows GDP Per Capita trend (if any)
```{r initial-plot}
plot(GDPPerCap~year, data = gdp)
```
Let's make things a little more eye catching using ggplot...
```{r ggplot-plot-fancy}
library(ggplot2)
ggplot(gdp, aes(x=gdp$year, y=gdp$GDPPerCap)) +
        geom_point(
                color="orange",
                fill="#69b3a2",
                alpha=0.5,
                size=2,
                shape=21,
                stroke=1
        ) +
        labs(title="GDP Per Capita NY, USA 1960-2016", x="Year", y="GDP Per Capita")
```
What these plots show us is that the data is _non-stationary_.  We therefore need to do some 'diffing' towards producing a stationary dataset.

# Stationary Data Assumption

* The mean and variance of a time series are constant for the whole series.

# Diffing

* The process of subtracting one observation from another.
* Used for transforming non-stationary data into stationary data.
* Example

X = [5,4,6,7,9,12,12]

What should the values of X after diffing it once be (ie., X^(1)^)?

X^(1)^ = [1,-2,-1,-2,-3,0]

What should be the values of X after diffing it twice (ie, X^(2)^)?

X^(2)^ = [1,-1,1,1,-3] # this is my answer so I hope it's right!

Diffing continues until any evident trend in the data is _removed_.

First we need to transform the dataframe into a time series dataset.
```{r create-timeseries-dataset}
us <- ts(gdp$GDPPerCap, start=min(gdp$year), end=max(gdp$year))
# install.packages("forecast")
library(forecast)
usOptimal <- auto.arima(us)
usOptimal
```
## Output of auto.arima() function explained - ie, ARIMA(p,d,q)

* **p** = the number of auto-regressive (AR) terms.
* **d** = the number of non-seasonal differences needed for stationarity.
* **q** = the number of lagged forecast errors in the prediction equation.

## What are the coefficients for our model?
```{r find-coefficients}
coef(usOptimal)
```
Coefficients ar1 and ar2 are the values of Beta1 ($\beta_1$) in our ARMA model.
Coefficients ma1 and ma2 are the values of Alpha1 ($\alpha_1$) in our ARMA model.


